---
title: "hw1"
author: "Yi-Ting Tsai"
format: html
editor: visual
---

This is the link to my repo: https://github.com/ytting920410/Homework1

import library

```{r}
library(readODS)
```

Q1:

a\.

```{r}
soybean <- read.csv('data/data.csv', header = TRUE)
class(soybean)
head(soybean)
summary(soybean)
colSums(is.na(soybean))
```

No NA in this dataset.

Adjust the data

```{r}
soybean$Season <- factor(soybean$Season)
soybean$Cultivar <- factor(soybean$Cultivar)

soybean$PH <- as.numeric(soybean$PH)
soybean$GY <- as.numeric(soybean$GY)
soybean$MHG <- as.numeric(soybean$MHG)
```

b\.

```{r}
nrow(soybean) # Number of observations
length(unique(soybean$Cultivar)) # Number of distinct cultivars
table_seasonal <- table(soybean$Season) # Showing the number of observation in each season
table_seasonal
```

There are 320 observations in the dataset and 40 distinct cultivars. There are 160 observations for each season.

c\.

```{r}
cor(soybean$PH, soybean$GY, method = 'pearson') # 1. correlation between PH and GY

cor_by_season <- sapply(
  split(soybean, soybean$Season), 
  function(x) cor(x$PH, x$GY, method = 'pearson')
)
cor_by_season # 2. correlation between PH and GY for each season

tallest <- soybean[soybean$PH == max(soybean$PH), ]
tallest$GY # 3. Find GY of the tallest observation

mean(soybean$MHG > mean(soybean$MHG)) * 100 # 4. Percentage of observations have MHG greater than avarage

cul_highGY <- aggregate(
  GY ~ Cultivar, 
  data = soybean, 
  FUN = function(x) mean(x)
)
cul_highGY[order(cul_highGY$GY, decreasing = TRUE), ][1, ] # 5. cultivar with the highest average GY
```

1.  The correlation is 0.1232807.
2.  Season 1 has the higher correlation.
3.  The grain yield of the observation with the greatest plant height is 3304.678.
4.  The percentage of observations with MHG greater than overall average MHG is 47.1875.
5.  Cultivar 98R30 CE has the highest average grain yield with 3964.968.

d\.

```{r}

```
