---
title: "hw1"
author: "Yi-Ting Tsai"
format: pdf
editor: visual
---

This is the link to my repo: https://github.com/ytting920410/STATS_506

## Import library
```{r}

```

## Create a function to check the input of every steps
```{r}
check_input <- function(n, p){
  if(length(n) != 1 || !is.numeric(n) || is.na(n) || 
     n < 1 || !is.finite(n) || n != floor(n)){
    stop('n must be a positive integer') # check n
  }
  if(length(p) != 1 || !is.numeric(p) || is.na(n) || 
     p < 0 || p > 1 || ! is.finite(p)){
    stop('p must be a real number between 0 and 1') # check p
  }
}
```

## a. Simulation function
```{r}
momentum_walk <- function(n, p){
  check_input(n, p) # sanitize the input
  steps <- matrix( # set possible moves
    c(1, 0,
      0, 1,
      -1, 0,
      0, -1), 
    ncol = 2, 
    byrow = TRUE
  )
  
  direction <- sample.int(4, size = 1) # the first step
  x <- steps[direction, 1] # the first step on the x axis
  y <- steps[direction, 2] # the first step on the y axis
  n <- n - 1 # we've done one step, so wee have n - 1 step left
  while(n > 0){ # n here is the number of remaining step we have
    if(runif(1) > p){ # for 1 - p chances we have to choose direction again
      direction <- sample.int(4, size = 1)
    }
    x <- x + steps[direction, 1] # update our position on x axis
    y <- y + steps[direction, 2] # update our position on y axis
    n <- n - 1 # n - 1 steps remain
  }
  c(x = x, y = y)
}
```

# Test the momentum_walk function
```{r}
momentum_walk(10, 0.6)
momentum_walk(100, 0.4)
```

## b. implement the simulation
# Version 1
Same as momentum_walk function above
```{r}
momentum_walk1 <- function(n, p){
  check_input(n, p) # sanitize the input
  steps <- matrix( # set possible moves
    c(1, 0,
      0, 1,
      -1, 0,
      0, -1), 
    ncol = 2, 
    byrow = TRUE
  )
  
  direction <- sample.int(4, size = 1) # the first step
  x <- steps[direction, 1] # the first step on the x axis
  y <- steps[direction, 2] # the first step on the y axis
  n <- n - 1 # we've done one step, so wee have n - 1 step left
  while(n > 0){ # n here is the number of remaining step we have
    if(runif(1) > p){ # for 1 - p chances we have to choose direction again
      direction <- sample.int(4, size = 1)
    }
    x <- x + steps[direction, 1] # update our position on x axis
    y <- y + steps[direction, 2] # update our position on y axis
    n <- n - 1 # n - 1 steps remain
  }
  c(x = x, y = y)
}
```

# Version 2
```{r}
momentum_walk2 <- function(n, p){
  check_input(n, p) # sanitize the input
  steps <- matrix( # set possible moves
    c(1, 0,
      0, 1,
      -1, 0,
      0, -1), 
    ncol = 2, 
    byrow = TRUE
  )
  
  direction <- sample.int(4, size = 1) # the first step
  all_direction <- c( # produce all n steps we will go
    direction, # add the first step
    if(n >= 2){ # if we have n = 1, we don't need further action
      lapply(2: n, function(i){ # for n >= 2, we have to calculate steps
        if(runif(1) > p){ # for 1 - p chances we have to choose direction again
          direction <<- sample.int(4, size = 1)
        }
        return(direction) # get next step
      })
    }else{
      integer(0) # no action
    }
  )
  final_pos <- colSums(steps[unlist(all_direction), , drop = FALSE])
  # calculate our final position by adding all the moves we made
  c(x = final_pos[1], y = final_pos[2])
}
```

# Test the momentum_walk2 function
```{r}
momentum_walk2(1, 0.6)
momentum_walk2(100, 0.4)
```

## c. Reproducibility
```{r}
set.seed(128)
result1_f1 <- momentum_walk1(10, 0.6)
set.seed(128)
result2_f1 <- momentum_walk1(10, 0.6)
set.seed(128)
result1_f2 <- momentum_walk1(1000, 0.6)
set.seed(128)
result2_f2 <- momentum_walk1(1000, 0.6)

result1_f1 == result2_f1
result1_f2 == result2_f2
```
set.seed sets the random number generator to a fixed initial state. As long as the seed, functions used are called in the same order, the same result will be reproduced.
