---
title: "hw2"
author: "Yi-Ting Tsai"
format: pdf
editor: visual
---

This is the link to my repo: https://github.com/ytting920410/STATS_506

## Import library

```{r}
library(microbenchmark)
library(roxygen2)
```

### Problem 1: Random Walk with Momentum

## Create a function to check the input of every steps

```{r}
#' Sanitize inputs for a momentum random walk
#'
#' Checks whether the inputs for a momentum random walk satisfy the required
#' conditions.
#'
#' @param n A positive integer specifying the number of steps.
#' @param p A numeric value between 0 and 1.
#'
#' @return This function returns \code{NULL} invisibly if both inputs are
#'   valid. Otherwise, it stops with an error message.
#'
#' @keywords internal
check_input <- function(n, p){
  if(length(n) != 1 || !is.numeric(n) || is.na(n) || 
     n < 1 || !is.finite(n) || n != floor(n)){
    stop('n must be a positive integer') # check n
  }
  if(length(p) != 1 || !is.numeric(p) || is.na(n) || 
     p < 0 || p > 1 || ! is.finite(p)){
    stop('p must be a real number between 0 and 1') # check p
  }
}
```

## a. Simulation function

```{r}
#' Simulate a momentum random walk
#'
#' Simulates an \code{n}-step two-dimensional random walk. The first
#' direction is chosen randomly. Each later step keeps the previous direction
#' with probability \code{p}; otherwise, a new direction is chosen randomly.
#'
#' @param n A positive integer indicating the number of steps.
#' @param p A number between 0 and 1 indicating the momentum probability.
#'
#' @return A named vector containing the final x and y coordinates.
#'
#' @examples
#' set.seed(123)
#' momentum_walk(10, 0.6)
momentum_walk <- function(n, p){
  check_input(n, p) # sanitize the input
  steps <- matrix( # set possible moves
    c(1, 0,
      0, 1,
      -1, 0,
      0, -1), 
    ncol = 2, 
    byrow = TRUE
  )
  
  direction <- sample.int(4, size = 1) # the first step
  x <- steps[direction, 1] # the first step on the x axis
  y <- steps[direction, 2] # the first step on the y axis
  n <- n - 1 # we've done one step, so wee have n - 1 step left
  while(n > 0){ # n here is the number of remaining step we have
    if(runif(1) > p){ # for 1 - p chances we have to choose direction again
      direction <- sample.int(4, size = 1)
    }
    x <- x + steps[direction, 1] # update our position on x axis
    y <- y + steps[direction, 2] # update our position on y axis
    n <- n - 1 # n - 1 steps remain
  }
  c(x = x, y = y)
}
```

# Test the momentum_walk function

```{r}
set.seed(44)
momentum_walk(10, 0.6)
momentum_walk(100, 0.4)
```

## b. implement the simulation

# Version 1

Same as momentum_walk function above

```{r}
#' Simulate a momentum random walk
#'
#' Simulates an \code{n}-step two-dimensional random walk with loop. The first
#' direction is chosen randomly. Each later step keeps the previous direction
#' with probability \code{p}; otherwise, a new direction is chosen randomly.
#'
#' @param n A positive integer indicating the number of steps.
#' @param p A number between 0 and 1 indicating the momentum probability.
#'
#' @return A named vector containing the final x and y coordinates.
#'
#' @examples
#' set.seed(123)
#' momentum_walk(10, 0.6)
momentum_walk1 <- function(n, p){
  check_input(n, p) # sanitize the input
  steps <- matrix( # set possible moves
    c(1, 0,
      0, 1,
      -1, 0,
      0, -1), 
    ncol = 2, 
    byrow = TRUE
  )
  
  direction <- sample.int(4, size = 1) # the first step
  x <- steps[direction, 1] # the first step on the x axis
  y <- steps[direction, 2] # the first step on the y axis
  n <- n - 1 # we've done one step, so wee have n - 1 step left
  while(n > 0){ # n here is the number of remaining step we have
    if(runif(1) > p){ # for 1 - p chances we have to choose direction again
      direction <- sample.int(4, size = 1)
    }
    x <- x + steps[direction, 1] # update our position on x axis
    y <- y + steps[direction, 2] # update our position on y axis
    n <- n - 1 # n - 1 steps remain
  }
  c(x = x, y = y)
}
```

# Version 2

```{r}
#' Simulate a momentum random walk
#'
#' Simulates an \code{n}-step two-dimensional random walk with lapply. The first
#' direction is chosen randomly. Each later step keeps the previous direction
#' with probability \code{p}; otherwise, a new direction is chosen randomly.
#'
#' @param n A positive integer indicating the number of steps.
#' @param p A number between 0 and 1 indicating the momentum probability.
#'
#' @return A named vector containing the final x and y coordinates.
#'
#' @examples
#' set.seed(123)
#' momentum_walk(10, 0.6)
momentum_walk2 <- function(n, p){
  check_input(n, p) # sanitize the input
  steps <- matrix( # set possible moves
    c(1, 0,
      0, 1,
      -1, 0,
      0, -1), 
    ncol = 2, 
    byrow = TRUE
  )
  
  direction <- sample.int(4, size = 1) # the first step
  all_direction <- c( # produce all n steps we will go
    direction, # add the first step
    if(n >= 2){ # if we have n = 1, we don't need further action
      lapply(2: n, function(i){ # for n >= 2, we have to calculate steps
        if(runif(1) > p){ # for 1 - p chances we have to choose direction again
          direction <<- sample.int(4, size = 1)
        }
        return(direction) # get next step
      })
    }else{
      integer(0) # no action
    }
  )
  final_pos <- colSums(steps[unlist(all_direction), , drop = FALSE])
  # calculate our final position by adding all the moves we made
  c(x = final_pos[1], y = final_pos[2])
}
```

# Test the momentum_walk2 function

```{r}
set.seed(44)
momentum_walk2(1, 0.6)
momentum_walk2(100, 0.4)
```

## c. Reproducibility

```{r}
set.seed(128)
result1_f1 <- momentum_walk1(10, 0.6)
set.seed(128)
result2_f1 <- momentum_walk1(10, 0.6)
set.seed(128)
result1_f2 <- momentum_walk1(1000, 0.6)
set.seed(128)
result2_f2 <- momentum_walk1(1000, 0.6)

result1_f1 == result2_f1 # test if these results are the same
result1_f2 == result2_f2 # test if these results are the same
```

set.seed sets the random number generator to a fixed initial state. As long as the seed, functions used are called in the same order, the same result will be reproduced.

## d. Does momentum mater?

```{r}
set.seed(32)
resultsD <- sapply(c(0, 0.5, 0.9), function(p) momentum_walk1(1000, p))
# run momentum walk 1000 times for p = 0, 0.5, 0.9
distance <- apply(t(resultsD), 1, function(pos) sqrt(sum(pos ^ 2)))
# calculate each distances from (0, 0) to (x, y)
resultsDdf <- data.frame(
  p = c(0, 0.5, 0.9), 
  x = resultsD['x', ], 
  y = resultsD['y', ], 
  distance = distance
)
resultsDdf
```

Increasing p tends to increase the final distance. Having higher p, the simulation tends to move toward the same way.

## e. Monte Carlo simulation

```{r}
p <- c(0, 0.25, 0.5, 0.75, 0.9) # for each of p
mean_distance <- sapply(
  p, 
  function(prob){ # do momentum walk 1000 times
    distance <- replicate(
      1000, 
      {
        position <- momentum_walk1(1000, prob) # do momentum walk with 1000 steps
        sqrt(sum(position ^ 2)) # calculate distance
      }
    )
    mean(distance) # get the mean of 1000 tries
  }
)
result_md <- data.frame(
  p = p, 
  distance = mean_distance
)
result_md
```

The mean of final distance increases as p increases.

## f. Compare computational performance

```{r}
BM_1000 <- microbenchmark( # test both methods with 1000 times
  loop = momentum_walk1(1000, 0.6), 
  reduce = momentum_walk2(1000, 0.6), 
  time = 100L
)
BM_100000 <- microbenchmark( # test both methods with 1000 times
  loop = momentum_walk1(100000, 0.6), 
  reduce = momentum_walk2(100000, 0.6), 
  time = 20L
)
print(BM_1000)
print(BM_100000)
```

i.  It seems version 1 is faster.
ii. The comparison result is the same as the smaller input.
iii. Version 2 has additional function calls and version 1 just updates the direction and position directly. So version 1 run faster.

### Problem 2: Monte Carlo Simulation of Campus Bus Arrivals

```{r}
set.seed(256)

n <- 10000 # 10000 days

time68 <- matrix( # simulated number of passengers in each hours from 6AM to 8AM
  rpois(n * 3, lambda = 4), 
  nrow = n, 
  ncol = 3
)

time9 <- matrix( # simulated number of passengers in 9AM
  rnorm(n, mean = 45, sd = 4), 
  nrow = n, 
  ncol = 1
)

time1015 <- matrix( # simulated number of passengers in each hours from 10AM to 3PM
  rpois(n * 6, lambda = 15), 
  nrow = n, 
  ncol = 6
)

time16 <- matrix( # simulated number of passengers in 4PM
  rnorm(n, mean = 55, sd = 5), 
  nrow = n, 
  ncol = 1
)

time1722 <- matrix( # simulated number of passengers in each hours from 5PM to 10PM
  rpois(n * 6, lambda = 9), 
  nrow = n, 
  ncol = 1
)

total <- rowSums(time68) + time9 + rowSums(time1015) + time16 + rowSums(time1722)
# sum of each day
mean(total) # calculate mean of 10000 days of data 
```

The simulated estimation is about 211 passengers per day.

### Problem 3: Linear Regression

## a. Data preparation and de-identification

```{r}
youtube <- read.csv( # read data
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv"
)

youtube_deiden <- youtube[!(names(youtube) %in% c(
  'brand', 
  'title',
  'link',
  'youtube_link',
  'channel_title',
  'category_id',
  'published_at',
  'thumbnail',
  'description'
))] # remove the columns we do not need
dim(youtube_deiden)
names(youtube_deiden)
```

There are 247 rows and 18 columns

## b. Exploring potential outcomes

```{r}
outcome <- c(
  'view_count', 
  'like_count', 
  'dislike_count', 
  'favorite_count', 
  'comment_count'
)

par(mfrow = c(2, 3)) # draw the distribution and print the count of favorite_count
for(i in outcome){
  hist(youtube_deiden[[i]],
    main = i,
    xlab = i, 
    breaks = 30
  )
}
table(youtube_deiden$favorite_count, useNA = 'ifany')
```

We can see the data are strongly right-skewed expect favorite_count. Thus, we can apply log(1 + x). We are not going to use favorite_count because most of it are 0.

```{r}
youtube_deiden$log_view_count <- log1p(youtube_deiden$view_count)
youtube_deiden$log_like_count <- log1p(youtube_deiden$like_count)
youtube_deiden$log_dislike_count <- log1p(youtube_deiden$dislike_count)
youtube_deiden$log_comment_count <- log1p(youtube_deiden$comment_count)
# calculate log(1 + x)

outcome_transformed <- c(
  'log_view_count', 
  'log_like_count', 
  'log_dislike_count', 
  'log_comment_count'
)

par(mfrow = c(2, 2))
for(i in outcome_transformed){ # draw the distribution
  hist(
    youtube_deiden[[i]], 
    main = i, 
    xlab = i, 
    breaks = 30
  )
}
par(mfrow = c(1, 1))
```

The transformed variables are used as the outcomes in the analyses. favorite_count is not suitable because is has not variation.

## c. Fitting linear regression models

```{r}
predictor <- c(
  "funny", 
  "show_product_quickly", 
  "patriotic", 
  "celebrity", 
  "danger", 
  "animals", 
  "use_sex"
)
# set the characteristics we want to use

to_int <- function(x){ # change those Boolean expression into integer
  if(is.logical(x)){
    return(as.integer(x))
  }
}

youtube_deiden[predictor] <-
  lapply(
    youtube_deiden[predictor],
    to_int
  )
# Transform logical predictors into 0 and 1
```

```{r}
models <- expand.grid(
  outcome = outcome_transformed, 
  predictors = predictor, 
  stringsAsFactors = FALSE
)

result_model <- do.call(
  rbind, 
  lapply(seq_len(nrow(models)), function(i){
      outcome_i <- models$outcome[i]
      predictors_i <- models$predictors[i]
      # select outcome and predictor for row i
      
      formula_i <- as.formula(
        paste(
          outcome_i, 
          "~", 
          predictors_i, 
          "+ year"
        )
      ) # construct formula
      
      fit_i <- lm(
        formula = formula_i, 
        data = youtube_deiden
      ) # fit the linear model
      
      # get the coefficient table from summary
      coef_table <- summary(fit_i)$coefficients
      
      data.frame( # form a data.frame with outcome, predictor, estimate, p_value
        outcome = outcome_i, 
        predictor = predictors_i, 
        estimate <- coef_table[predictors_i, "Estimate"], 
        p_value <- coef_table[predictors_i, "Pr(>|t|)"], 
        stringsAsFactors = FALSE
      )
    }
  )
)

names(result_model) <- c(
  "outcome", "predictor", "estimate", "p_value"
)

result_model$significant <- result_model$p_value < 0.05
result_model$direction <- ifelse(result_model$estimate > 0, "positive", "negative")
# add significant and direction
result_model
```

1.  None of the commercial characteristics are statistically significant.
2.  All of them are negative.

```{r}
result_view_ct <- subset(result_model, outcome == "log_view_count")
largest_view_ct <- result_view_ct[which.max(result_view_ct$estimate), 
                                  c("outcome", 
                                    "predictor", 
                                    "estimate", 
                                    "p_value", 
                                    "significant", 
                                    "direction")
                                  ]
# get the row with outcome is log_view_count and largest estimate value
largest_view_ct
```

3.  Commercial characteristic danger has the largest positive estimated association with view count

## d. Estimating regression coefficients using matrix algebra

```{r}
result_view <- subset(
  result_model, 
  outcome == "log_view_count"
)
# get regression results for log_view_count only
result_view
```

```{r}
selected_pred <- result_view$predictor[which.min(result_view$p_value)]
# get the commercial characteristic with the smallest p-value
selected_pred
```

```{r}
variables <- c("log_view_count", selected_pred, "year")
# choose the variables we need
data_ols <- youtube_deiden[
  complete.cases(youtube_deiden[variables]), 
  variables, 
  drop = FALSE
] # keep only these variables
head(data_ols)
```

```{r}
X <- cbind(
  intercept = 1, 
  predictor = data_ols[[selected_pred]], 
  year = data_ols$year
)
# construct X matrix in OLS

y <- data_ols$log_view_count

dim(X)
head(y)
length(y)
# construct y in OLS
```

```{r}
# build beta_hat
beta_hat <- solve(t(X) %*% X) %*% (t(X) %*% y)
rownames(beta_hat) <- c(
  "(Intercept)", 
  selected_pred, 
  "year"
)

beta_hat
# print beta_hat

formula_OLS <- as.formula( # construct the formula
  paste(
    "log_view_count ~", 
    selected_pred, 
    "+ year"
  )
)
fit_OLS <- lm(formula = formula_OLS, data = data_ols)
# fit the formula with function lm()

coef(fit_OLS)
# get beta_hat from lm()
```

```{r}
all.equal(as.numeric(beta_hat), as.numeric(coef(fit_OLS)))
# test if our beta_hat is equal to the one from lm()
```

The result of testing if $\hat{\beta}$ and coefficient from lm() is equal is TRUE
